Do motif roles provide unique information about species risk of extinction?\\ \medskip Supplementary Information

### 2021-10-06 manuscript/reorder_main_text.tex
Updates from Overleaf

    % Anna AA: The whole article is now approximately 4400 words (not counting fig texts); intro 850, methods 1600, results 620, discussion 1300. See discussion on section statistical analysis in the methods, I think we can make it shorter
    % - glossary
    % - refs
    % - ending

    % - Fig. 4 not much in discussion, could be a candidate for the SI? Or add more discussion about apparent competition, etc. in discussion sections per motif

@@

	\subsection*{Calculating persistence}	
    % basic idea behind BN modeling
        Bayesian Networks can be used to represent species and their feeding interactions as the nodes and links of a network (see Appendix \emph{S2} for full details).
        In our setting, each species (node) is a Bernoulli random variable, being either extant or extinct. 
        The probability that a particular species is extinct is a function of the states of its resources.

@@

    There are limitations to the approach presented here.
    Most importantly, the Bayesian Network framework operates on a  strict bottom-up principle and therefore cannot include top-down effects in the extinction risk calculations~\citep{Eklof2013}. 
    Here our focus was to analyze the importance of disturbances to the basal resources, in which case the Bayesian network approach captures a majority of the secondary extinctions obtained in a fully-fledged dynamical model with much greater efficiency~\citep{Eklof2013}.
    If top-down effects are important for a particular empirical system, however, our simulations may not be accurate. 
    Ongoing work using a dynamical model shows that time to extinction for species following disturbance is tightly coupled to the motifs a species participates in, and that adding top-down effects can lead omnivory to be negatively correlated with persistence \citep{Cirtwill2021_inprep}. 
    If top-down effects are not crucial, the Bayesian Network approach could be extended by explicitly incorporating information on species traits.
    In empirical webs, species traits are related to both motif participation~\citep{cirtwill2018feeding} and extinction risk~\citep{Brose2017, curtsdotter2011robustness, Cardillo2005, Purvis2000},
    adding an extra dimension of information which we do not consider here.

@@

    \begin{figure}
        \centering
        \includegraphics[width=\textwidth]{figures/roles_vs_TL.eps}
        \caption{As with motifs, persistence varied with simpler measures of species roles in networks. Moreover, motif frequencies correlated with in-degree and trophic level. \textbf{A-B)} Persistence increased with increasing in-degree for low levels of disturbance to basal resources but decreased with increasing degree for high levels of disturbance to basal resources.
        Persistence declined with inceasing STL for all levels of disturbance to basal resources.
        \textbf{C-D)} The proportion of omnivory increased with increasing degree while all other proportions decreased. The proportions of omnivory and direct competition decreased with increasing STL while the proportions of apparent competition and three-species chains increased.}
        \label{fig:motifs_vs_TL_and_deg}
    \end{figure}

>>>>

@@

	\subsection*{Calculating persistence}	
    % basic idea behind BN modeling
        Bayesian networks can be used to represent species and their feeding interactions as the nodes and links of a network (see Appendix \emph{S2} for full details).
        In our setting, each species (node) is a Bernoulli random variable, being either extant or extinct. 
        The probability that a particular species is extinct is a function of the states of its resources.

@@

    There are limitations to the approach presented here.
    Most importantly, the Bayesian network framework operates on a  strict bottom-up principle and therefore cannot include top-down effects in the extinction risk calculations~\citep{Eklof2013}. 
    Here our focus was to analyze the importance of disturbances to the basal resources, in which case the Bayesian network approach captures a majority of the secondary extinctions obtained in a fully-fledged dynamical model with much greater efficiency~\citep{Eklof2013}.
    If top-down effects are important for a particular empirical system, however, our simulations may not be accurate. 
    Ongoing work using a dynamical model shows that time to extinction for species following disturbance is tightly coupled to the motifs a species participates in, and that adding top-down effects can lead omnivory to be negatively correlated with persistence \citep{Cirtwill2021_inprep}. 
    If top-down effects are not crucial, the Bayesian network approach could be extended by explicitly incorporating information on species traits.
    In empirical webs, species traits are related to both motif participation~\citep{cirtwill2018feeding} and extinction risk~\citep{Brose2017, curtsdotter2011robustness, Cardillo2005, Purvis2000},
    adding an extra dimension of information which we do not consider here.

@@

    \begin{figure}
        \centering
        \includegraphics[width=\textwidth]{figures/roles_vs_TL.eps}
        \caption{Motif participation correlated with in-degree and trophic level, and both of these simpler role measures were also correlated with persistence. \textbf{A-B)} The proportion of omnivory increased with increasing degree while all other proportions decreased. The proportions of omnivory and direct competition decreased with increasing STL while the proportions of apparent competition and three-species chains increased. \textbf{C-D)} Persistence increased with increasing in-degree for low levels of disturbance to basal resources but decreased with increasing degree for high levels of disturbance to basal resources.
        Persistence declined with inceasing STL for all levels of disturbance to basal resources.}
        \label{fig:motifs_vs_TL_and_deg}
    \end{figure}

### 2021-10-06 manuscript/supplement_motifs.tex
Updates from Overleaf

        We use Bayesian Network modelling to calculate the persistence of each species in our simulated networks. 
        A Bayesian Network represents a graphical model for probabilistic relationships among a set of variables. The nodes of the network represents the variables, and the links represent conditional dependencies among the variables. Each node has a set of local probability dependencies, conditional on the state of the parent nodes. 
        In our ecological setting, species act as nodes and are Bernoulli random variables, being either extant or extinct. The links act as feeding interactions between the nodes.
        The extinction probability of each species is a function of the state of its resources.

>>>>

        We use Bayesian network modelling to calculate the persistence of each species in our simulated networks. 
        A Bayesian network represents a graphical model for probabilistic relationships among a set of variables. The nodes of the network represents the variables, and the links represent conditional dependencies among the variables. Each node has a set of local probability dependencies, conditional on the state of the parent nodes. 
        In our ecological setting, species act as nodes and are Bernoulli random variables, being either extant or extinct. The links act as feeding interactions between the nodes.
        The extinction probability of each species is a function of the state of its resources.
